---
title: "Homework 1"
author: "Wenhao Gou - wg2364"
date: "2020.9.15"
output: github_document
---
##1. Overview and preparation:
This is the solution of homework 1 for course P8105. To setup, we need to import package `tidyvers` and `palmerpenguins`
```{r Include_packages, message = FALSE}
#Include packages
library(tidyverse)
library(palmerpenguins)
```

##2.Solution of Question 1
###2.1 Overview of the quesiton
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

###2.2 Create the dataframe
Use function `tibble`, we can create a data frame with random variables, numbers, logical vectors and factor vectors. We can view this dataframe use `solution_df`to see the dataframe and the  data type of each vector. (`dbl` indicate "double", `lgl` indicate logic, `chr` indicate character, and `fct` indicate factor)
```{r Df creation and view, collapse = TRUE}
#Create the dataframe:
solution_df <- tibble(
  vec_sample = rnorm(10) , 
  vec_logical = vec_sample > 0 ,
  vec_char = LETTERS [1:10] ,
  vec_factor = factor(c("A","B","C","C","B","A","A","B","C","C"))
)

#View the dataframe
solution_df
```
Also, we can try to take mean of each of subsets:
```{r Take_mean, collapse = TRUE}
#Take mean of each vector (we can also ues function "mean(pull(solution_df,col_you_want))" )
mean(solution_df$vec_sample)
mean(solution_df$vec_logical)
mean(solution_df$vec_char)
mean(solution_df$vec_factor)
```
This indicate that we can only do numeric calcuation on numbers and logic values. Also, we can deduce from the result that in the logic vectro, `TURE` equals to 1 and `FALSE` equal to 0. ( ``r length(solution_df$vec_sample[solution_df$vec_sample>0])`` numbers in `vec_sample` are greater than 0 and ``r length(solution_df$vec_sample[solution_df$vec_sample<=0])`` numbers in `vec_sample` are not greater than 0)

For characters and factors, the function `mean` will return `NA` (Not a number). This indicate that we cannot take mean or other numeric calculation on them

###2.3 Data type conversion:
In this part, we try to do these trnaformations: 

* Convert the logical vector to numeric, and multiply the random sample by the result

* Convert the logical vector to a factor, and multiply the random sample by the result

* Convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
```{r Transformation, results = 'hide', warning = FALSE}
#All the results were hided
#1 Logic -> Numeric, this part can operate with no error.
logic_to_numeric = as.numeric(solution_df$vec_logical)
logic_to_numeric * solution_df$vec_sample

#2 Logic -> Factor, error occured in line #2, but not in line #1
logic_to_factor = as.factor(solution_df$vec_logical)
logic_to_factor * solution_df$vec_sample

#3 Logic -> Factor -> Numeric, this part can operate with no error. 
factor_to_numeric = as.numeric(logic_to_factor)
factor_to_numeric * solution_df$vec_sample
```
The result can tell us more about the characteristic of these three data type and how they were converted within each function:

* Calculation result after the first transformation is:
``r logic_to_numeric * solution_df$vec_sample``. In this transformation, R converted Logic vector to numeric term in this manner: `TURE` = 1 and `FALSE` = 0. For example, ``r (solution_df$vec_sample[solution_df$vec_sample>0])[1]``, position ``r which(solution_df$vec_sample>0)[1]``  in the `vec_sample` remain the same in this result, while ``r (solution_df$vec_sample[solution_df$vec_sample<=0])[1]``, position ``r which(solution_df$vec_sample<=0)[1]`` in the sample is equal to `0`in this result. 

* Calculation result after the second transformation is all `NA`. But in the first line of code in this trasformation executed succesfully, and have the result ``r logic_to_factor`` as a roll of facter, wich have two item `TRUE` and `FALSE`.Note that is not same as the logic term, which can be convert into 0 and 1. As proved in the calculation of mean, this term cannot do numeric calculations.

* Calculation result after the third transformation is ``r factor_to_numeric * solution_df$vec_sample``. In this transformation, R converted factor vector to numeric term in this manner: Factor_Level 1 = 1, Factor_Level 2 = 2 etc. The factor name are arragned alphabetically (increasing order, see `?factor` for detail). In this sample, level  `FALSE` equal to 1 and level `TRUE` equal to 2. 
In this example, ``r (solution_df$vec_sample[solution_df$vec_sample<=0])[1]``
, position ``r which(solution_df$vec_sample<=0)[1]``  in the `vec_sample` is ``r (factor_to_numeric * solution_df$vec_sample)[which(solution_df$vec_sample<=0)[1]]`` (`FALSE = 1`)
 ``r (solution_df$vec_sample[solution_df$vec_sample>0])[1]``, position ``r which(solution_df$vec_sample>0)[1]`` in the `vec_sample` is ``r (factor_to_numeric * solution_df$vec_sample)[which(solution_df$vec_sample>0)[1]]``(`TRUE = 2`).  

##Solution of Question 2
###3.1 Overview of the question
This problem focuses the use of inline R code, plotting, and the behavior of `ggplot` for variables of different types.

###3.2 Load the dataset
Firstly, we need to load the dataset and have an overview about it. The first line of the code will load two dataset in the package, `penguins` and `penguins_raw` into the environment directly.In this quesiton, we will use dataset `penguins` only.
```{r Read and inspect the data}
#Load the dataset
data("penguins", package = "palmerpenguins")
```
###3.2 Description of the dataset
We can use `str` function to see an overview of this dataset.  We can also type `?penguins` in the console for a detailed description of all the variables. 
 This dataset have ``r nrow(penguins)`` rows and ``r ncol(penguins)`` colomus. In this dataset, we have 8 variables ``r names(penguins) ``. In these variables, `species`,`island ` are factors with 3 levels; `bill_length_mm`, `bill_depth_mm`,`flipper_length_mm`,`body_mass_g` are numbers denoting bill length(millimeters), bill depth(millimeters), flipper length (millimeters), and body mass(grams), have means `r mean(penguins$bill_length_mm)`,`r mean(penguins$bill_length_mm)`,`r mean(penguins$bill_length_mm)`,`r mean(penguins$bill_length_mm)`; `sex` is a factor denote the sex of the penguin and `year` is an integer range from 2007 to 2009. Missing value of varible `flipper_length_mm` and `penguins$bill_length_mm` are in line ``r which(is.na(penguins$bill_length_mm))[1]`` and ``r which(is.na(penguins$bill_length_mm))[2]``. 
 
###3.3 Plotting and saving
Next, we make a scatterplot of `flipper_length_mm` (y) vs `bill_length_mm` (x); color points using the `species` variable. Finally, we use `ggsave` function to save the result in the working directory. 
```{r Plotting and saving, collapse = TRUE, message = FALSE, warning = FALSE}
#Plotting
Penguins_Scatter <- penguins %>% 
  ggplot(aes(x=bill_length_mm,y=flipper_length_mm, color = species)) + 
  geom_point()

#Display the plot
Penguins_Scatter
#Saving the plot
ggsave('Penguins_Scatter.png', plot = Penguins_Scatter)
```
